Pop-up cycleways and pedestrian spaces: a peak into post Covid-19 transport futures

### 2020-05-09 conversation.Rmd
edits

In the coming weeks and months, active travel will play a vital role in enabling the reopening of many countries. As travel increases, we must ensure that safe, socially distanced walking and cycling remains possible. Maintaining social distancing on public transport could severely reduce the capacity of rail and bus networks, and there is a risk that this may lead to a rapid rise in road traffic. Without careful management, increased congestion could deter others from the streets, especially as [ record bicycle sales ] suggest that many people have taken advantage of lockdown conditions to take up or return to cycling. These people may have less experience in dealing with high volumes of road traffic.

@@

Local and national governments worldwide are still working out how best to respond to these challenges. On Saturday 9th May, British Transport Secretary Grant Shapps announced a £2 billion investment in measures to promote walking and cycling. This follows a £10 million commitment from the Scottish Government and action from Wales. Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan] to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf). Never has there been such a need for affordable, healthy and socially distanced transport that does not create the fumes that [emerging evidence] links to coronavirus susceptibility.

@@

With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised. As with the medical science, research is needed now. Methods developed to identify locations of high walking and cycling potential can help ensure that the ‘pop-up’ infrastructure that goes in is effective, safe, and placed where it is most needed. Agile research can respond to rapidly changing conditions, as government priorities shift and different sectors of the economy reopen.

@@

The methods have focused on road capacity and cycling potential, identifying roads where a ‘spare lane’ could be repurposed for cycling. Cycle potential is assessed based on travel to work data and estimates of cycling uptake under the UK Government’s national target of doubling cycling levels by 2025, with the length of road segments also being used to filter the top schemes.

The use of open data and publicly accessible models is a key aspect of the research. Given the global nature of the challenges we face, open and collaborative research is vital so that methods can be applied by as wide a user base as possible. The potential for international application can be seen for example in [ WHO report ] modelling cycling uptake in cities in low income countries. Within the UK, [ 50+ Local Authorities ] are now using the PCT to inform their strategic cycle network plans.

Maps (see below) show the recommendations for where to introduce pop-up cycleways. With the situation rapidly developing, some of the roads shown here, such as [ Princess Street in Manchester ], have already been identified by Local Authorities as targets for intervention.

>>>>

In the coming weeks and months, active travel will play a vital role in enabling the reopening of many countries. As travel increases, we must ensure that safe, socially distanced walking and cycling remains possible. Maintaining social distancing on public transport could severely reduce the capacity of rail and bus networks, and there is a risk that this may lead to a rapid rise in road traffic. Without careful management, increased congestion could deter others from the streets, especially as [ record bicycle sales ] suggest that many people have taken advantage of lockdown conditions to take up or return to cycling. Some of these cyclists may have less confidence in dealing with high volumes of road traffic.

@@

Local and national governments worldwide are still working out how best to respond to these challenges. On Saturday 9th May, British Transport Secretary Grant Shapps announced a £2 billion investment in measures to promote walking and cycling. This follows a £10 million commitment from the Scottish Government and action announced in Wales. Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan] to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf). Never has there been such a need for affordable, healthy and socially distanced transport that does not create the fumes that [emerging evidence] links to coronavirus susceptibility.

@@

With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised. As with the medical science, research is needed now. Methods developed to identify locations of high walking and cycling potential can help ensure that the ‘pop-up’ infrastructure that goes in is effective, safe, and placed where it is most needed. Agile research can respond to rapidly changing conditions, as government priorities shift and different sectors of the economy reopen. Today the greatest need may be for medical workers to safely access [hospitals], tomorrow it could be access to schools or retail centres.

@@

The methods have focused on road capacity and cycling potential, identifying roads where a ‘spare lane’ could be re-purposed for cycling. Cycle potential is assessed based on travel to work data and cycling uptake scenarios under the UK Government’s national target of doubling cycling levels by 2025, with the length of road segments also being used to filter the top schemes.

The use of open data and publicly accessible models is a key aspect of this research. Given the global nature of the challenges we face, open and collaborative research is vital so that methods can be applied by as wide a user base as possible. The potential for international application can be seen for example in [ WHO report ] modelling cycling uptake in cities in low income countries. Within the UK, [ 50+ Local Authorities ] are now using the PCT to inform their strategic cycle network plans.

Maps (see below) show the recommendations for where to introduce pop-up cycleways. With the situation rapidly developing, some of the roads shown here, such as [Princess Street in Manchester](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175), have already been identified by Local Authorities as targets for intervention.
